%!TEX encoding = UTF-8 Unicode

%% Thesis cover pages
%%
%% Based on the example.tex file provided by Tomás Oliveira e Silva
%%
%% Modified by João Paulo Barraca <jpbarraca@ua.pt>
%% Modified by Filipe Manco <filipe.manco@ua.pt>

% Remove 'coverdepname' parameter to remove department name from cover pages
% Add 'final' parameter to remove "DOCUMENTO PROVISÓRIO" from cover page
\documentclass[11pt,twoside,a4paper,final]{report}

%%% This file is a template for the cover of your
%%%  thesis.
%%% Lines like this one, starting with %%% , are
%%%  instructions. Search them to know where and
%%%  how to edit this file. After editing the
%%%  document you can delete them.
%%% Items marked with OPTIONAL can be removed.
%%% Most of the lines you will have to change
%%%  already have examples to help you understand.
%%% The complete list of options and commands can
%%%  be found in uaThesisTemplate.sty


%%% Select your area of study and department from
%%%  the following lists and replace them in the
%%%  \usepackage declaration ([area,department]).
%%%
%%% | area        |   | department |
%%% ---------------   --------------
%%% | arts        |   | dao        |
%%% | sciences    |   | dbio       |
%%% | education   |   | dcspt      |
%%% | economy     |   | deca       |
%%% | engineering |   | degeit     |
%%% | humanities  |   | de         |
%%% | health      |   | deti       |
%%% |             |   | demac      |
%%% |             |   | decivil    |
%%% |             |   | dem        |
%%% |             |   | fis        |
%%% |             |   | dlc        |
%%% |             |   | dmat       |
%%% |             |   | dq         |
%%% |             |   | dcm        |
%%% |             |   | geo        |
%%% ---------------   --------------
%% Import the UA thesis template
\usepackage[engineering,deti]{uaThesisTemplate}


%% Optional packages for utf8 encoding and T1 fonts
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

%% For prettier pages
\usepackage{lmodern}

%% We will use english and portuguese
\usepackage[english,portuguese]{babel}

\usepackage{xspace}% used by \sigla

%%% This package is only used to create a dummy
%%%  text for the abstract. You can remove the
%%%  following line in the final document.
\usepackage{blindtext}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%       DEFINE THESIS COVER FIELDS
%%                  START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Edit thesis year
%% Thesis Year
\ThesisYear{\the\year}


%%% Uncomment one of the following
%\DocumentTypeThesis
\DocumentTypeDissertation

%%% Edit author's name
%% Author's full name
\Author{Fábio Miguel\\Correia Alves}

%%% Edit thesis title in Portuguese
%% Thesis Mandatory Title in Portuguese
\Title{\selectlanguage{portuguese}Braço Robótico e Humano Desempenham Tarefas de forma Colaborativa}

%%% OPTIONAL: you can write your thesis title in
%%%  other languages. You can repeat the following
%%%  line as many times as you want, or remove it
%%%  if you only want the portuguese title.
%% Optional Titles in additional languages.
\Title{\selectlanguage{english}Collaborative Tasks between Robotic Manipulator and Human}

%%% OPTIONAL: add a quotation if you like.
%% Optional Quotation
% \Quote{\selectlanguage{english}The greatest challenge to any thinker is stating
% the problem in a way that will allow a solution}{Bertrand Russell}

%%% Edit your supervisor's information.
%%%
%%% \Supervisor{name}{category}
%%%
%%% name   : Supervisor's name
%%% categoy: One of the following list
%%%
%%% | category               |
%%% --------------------------
%%% | catedrático            |
%%% | catedrático convidado  |
%%% | associado c/ agregação |
%%% | associado              |
%%% | auxiliar               |
%%% | auxiliar convidado     |
%%% --------------------------
%% Supervisor information
\Supervisor{José Nuno Panelas Nunes Lau}{associado}

%%% OPTIONAL: you can add a second supervisor.
%%%
%%% \SecondSupervisor{name}{category}
%%%  {[preposition: (at, in, no, da)] department}
%%%  {[preposition: (at, in, do, na)] school}
%%%
%%% name      : See above.
%%% category  : See above.
%%% department: Can be a department, faculty or
%%%             other; Choose the preposition
%%%             accordingly.
%%% school    : University/Scholl name; Choose the
%%%             preposition accordingly.
%% Optional SecondSuperviser information
\SecondSupervisor{Manuel Bernardo Salvador Cunha}{auxiliar}{do Departamento de Eletrónica, Telecomunicações e Informática}{da Universidade de Aveiro}

%%% Edit your degree information.
%%%
%%% \Degree{title}{name}
%%%
%%% title: Degree title (Mestre or Doutor)
%%% name : Degree name
%% Degree title and name
\Degree{Mestre}{Engenharia de Computadores e Telemática}

%%% OPTIONAL: edit grant support texts where
%%%  appropriate. Maximum two.
%%%
%%% \GrantText{grant1}{grant2}
%% Optional Grant support texts
% \GrantText{Texto Apoio financeiro do POCTI no âmbito do III Quadro Comunitário de Apoio.}
%           {Texto Apoio financeiro da FCT e do FSE no âmbito do III Quadro Comunitário de Apoio.}

%%% OPTIONAL: Add a dedication text.
%% Optional Dedication text
% \Dedication{Dedico este trabalho à minha esposa e filho pelo incansável apoio.}

%%% OPTIONAL: Add an acknowledgement text
%% Optional Acknowledgement text
\Acknowledgements{Agradeço ao Professor Nuno Lau e ao Professor Bernardo Cunha toda a disponibilidade e esclarecimentos ao longo do desenvolvimento deste trabalho. Não só a eles mas também a todos os membros do IRISLab, pelas oportunidades de aprendizagem e crescimento que me proporcionaram ao longo do meu percurso académico. \\
Ao meu parceiro de investigação Marcelo Fraga, tanto pelas longas e arduas discussões como por toda a camaradagem que criamos. Foi quem mais me provou que 2 cabeças pensam melhor que 1, e que quando eu quero, consigo ser incompreensivelmente teimoso. \\
À Margarida, que primeiro por acaso, e depois com muito gosto, me acompanhou desde o início deste percurso, e cujo conhecimento que me transmitiu, não só académico, foi imprescindível para me tornar na pessoa que sou hoje. \\
À Fernanda Rocha, que sem querer, juntou o melhor grupo de idiotas que alguma vez pisou esta terra. Foram 5 anos de amizade e maluqueiras, com algum estudo intensivo pelo meio, que irei levar para a vida. \\
Finalmente, um obrigado muito especial aos meus pais, pelo apoio incansável ao longo destes anos. Irei ter um trabalho muito árduo para vos recompensar por tudo o que me proporcionaram até agora.}

%%% Edit the keywords and abstract in Portuguese
%%% Don't forget to remove the \blindtext[3]
%%%  declaration.
%%% Don't remove the \selectlanguage statement or you will end with incorrect hyphenation
%% Mandatory Keywords and Abstract in Portuguese
\Abstract{Palavras Chave}{colaboração humano robô, robôs colaborativos, guiamento manual, transferência de ferramentas, prevensão de colisões, arquitetura de tarefas, controlo de movimento em tempo real}
         {Resumo}{\selectlanguage{portuguese}A colaboração humano robô tem-se tornado um tema proeminente em ambientes industriais ao longo dos anos. O aparecimento de robôs colaborativos e as vantagens que eles trazem para as linhas de montagem causam a necessidade de explorar seu hardware e criar novos sistemas que permitam a execução segura de tarefas colaborativas entre humanos e robôs. Essas vantagens incluem as suas reduzidas dimensões e custos, os seus materiais de construção mais leves, a existência hardware de assistência de força/precisão e o cumprimento de requisitos de segurança, que os permitem ser considerados como colaborativos. 
         Nesta Dissertação, propomos um conjunto de ferramentas e técnicas que possibilitam a execução de tarefas colaborativas entre um humano e um braço robótico. Essas técnicas incluem compensação de peso acoplado à garra do robô, identificação de obstáculos dinâmicos no ambiente e a criação de novas tarefas de forma simplificada e abstrata. Essas técnicas podem então ser unidas para criar várias tarefas colaborativas. Exemplos, implementados e testados, são a transferência de objetos entre humano e robô, manipulação da posição do robô através de interação direta com a sua garra, manipulação de objetos acoplados dinamicamente e a execução de uma tarefa industrial com prevenção de colisões. Todas essas tarefas foram agrupadas numa máquina de estados que permite ao utilizador interagir dinamicamente com o robô, mudar a tarefa em execução, enquanto recebe feedback constante.
         Estas tarefas foram implementadas e testadas num ambiente real com um Universal Robots UR10e. O cenário consiste num espaço de trabalho partilhado onde o humano e o robô podem interagir fisicamente e realizar as tarefas propostas.}

%%% OPTIONAL: you can write the keywords and
%%%  abstract in other languages. You can repeat
%%%  the following lines as many times as you want
%%%  or remove them if you only want portuguese.
%%% Don't remove the \selectlanguage statement or you will end with incorrect hyphenation.
%%% You can modify the language to any other supported by Latex
%% Optional Keywords and Abstract in additional languages.
\Abstract{Keywords}{human robot collaboration, collaborative robots, hand guiding, tool transfer, collision avoidance, task level architecture, real time motion control}
         {Abstract}{\selectlanguage{english}Human Robot Collaboration has become a prominent subject in industry settings over the years. The insurgence of collaborative robots and the advantages they bring to assembly lines have caused the need to exploit their hardware and create new systems that can enable the safe execution of collaborative tasks between humans and robots. Such advantages include their reduced size and cost, their lightweight construction, their force/precision assistance hardware and their safety requirements, that must be met in order to be considered collaborative.
         In this Dissertation, we propose a set of tools and techniques that enable the execution of collaborative tasks between a human and a robotic manipulator. Such techniques include dynamic payload compensation at the end effector of the robot, identification of moving obstacles in the environment, and the creation of high level task plans. This techniques can then be joined together to create various collaborative tasks. Examples of such, both implemented and tested, are the transfer of objects between human and robot, precise hand guiding of the robot at the end effector, hand guided manipulation of dynamically coupled objects, and the execution of an industrial task with collision avoidance. All of this tasks were grouped in a global state machine that allows the user to seamlessly interact with the robot, change the task in execution while receiving constant feedback.\\
         These tasks were implemented and tested in a real setting with an Universal Robots UR10e. The setup consisted on a shared workspace where the human and the robot could physically interact and perform the proposed tasks.
         }
% TODO: Completar o abstract, espetar aqui tudo o que foi feito

%%% Edit the Jury President information.
%% First Member will be the President of the Jury
\JuriMember{Professor Doutor José Alberto Gouveia Fonseca}{professor associado do Departamento de Eletrónica, Telecomunicações e Informática da Universidade de Aveiro}

%%% Repeat the above declaration for additional
%%%  members of the jury as many times as necessary.
%% Additional members of the Jury
\JuriMember{Professor Doutor Pedro Mariano Simões Neto}{professor auxiliar do Departamento de Engenharia Mecânica da Universidade de Coimbra}
\JuriMember{Professor Doutor José Nuno Panelas Nunes Lau}{professor associado do Departamento de Eletrónica, Telecomunicações e Informática da Universidade de Aveiro}
% \JuriMember{-----}{-----}
% \JuriMember{-----}{-----}
% \JuriMember{-----}{-----}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%       DEFINE THESIS COVER FIELDS
%%                  END
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{document}

%%% By default we consider portuguese to be the
%%%  main language (the last one in the usepackage
%%%  declaration). If you're writing your thesis
%%%  in english uncomment the following line.
%%% Set the proper language or you will end with incorrect hyphenation!
%\selectlanguage{english}

\MakeCover

\end{document}
